---
title: "Lesson 3: Introduction to Simulations"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#006a4e"
date: "`r library(here); source(here('class_dates.R')); w2d1`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: Lesson 3 Slides
    html-math-method: mathjax
---

```{r setup, include=FALSE}
library(here)
source(here("class_dates.R"))
```

# Learning Objectives

1. intro to simulations
2. discrete RVs simmies
3. continuous RVs simmies

## Recall: Outcomes, events, sample spaces

::: definition
::: def-title
Definition: Outcome
:::

::: def-cont
The possible results in a random phenomenon.
:::
:::

::: definition
::: def-title
Definition: Sample Space
:::

::: def-cont
The **sample space** $S$ is the set of *all* outcomes
:::
:::

::: definition
::: def-title
Definition: Event
:::

::: def-cont
An **event** is a *collection of some* outcomes. An event can include multiple outcomes or no outcomes (a subset of the sample space).
:::
:::

When thinking about events, think about outcomes that you might be asking the probability of. For example, what is the probability that you get a heads or a tails in one flip? (Answer: 1)

## What is a simulation?

A **probability model** for a random phenomenon includes a sample space, events, random variables, and a probability measure.

::: definition
::: def-title
Simulation
:::
::: def-cont
**Simulation** involves using a probability model to artificially recreate a random phenomenon, many times, usually using a computer.
:::
:::

We simulate outcomes and values of random variables according to the model's assumptions.

## The Foundation: Relative Frequencies
- Probabilities can be interpreted as **long-run relative frequencies**
- By simulating a random phenomenon a large number of times, we can approximate the probability of an event by calculating the relative frequency of its occurrence.
- Simulation is a powerful tool for approximating probabilities, distributions of random variables, long-run averages, and other characteristics.


## Still confused about long-run relative frequencies?

[Seeing Theory, Chapter 1: Basic Probability, Chance Events](https://seeing-theory.brown.edu/basic-probability/index.html)

```{css}
#| echo: false

<style>
#wrap { width: 600px; height: 390px; padding: 0; overflow: hidden; }
#frame { width: 1300px; height: 600px; border: 1px solid black; }
#frame { zoom: 1.5; -moz-transform: scale(1.5); -moz-transform-origin: 0 0; }
</style>
```

```{=html}
<iframe id="frame" src="https://seeing-theory.brown.edu/basic-probability/index.html"></iframe>
```


## 4 (S)teps of a Simulation

::: proposition
::: prop-title
1. Set up
:::
::: prop-cont
Define the probability space and related random variables and events, including assumptions. 
:::
:::

::: proof1
::: proof-title
2. Simulate
:::
::: proof-cont
Run the simulation to generate outcomes according to the assumptions.
:::
:::

::: example
::: ex-title
3. Summarize
:::
::: ex-cont
Analyze the output using plots and summary statistics like relative frequencies and averages.
:::
:::

::: extra
::: ext-title
4. Sensitivity analysis
:::
::: ext-cont
Investigate how results change when assumptions or parameters of the model are altered.
:::
:::

## Types of simulations

- We can simulate discrete and continuous random variables

 

- **Discrete random variables** (RVs) are a little easier to understand
  - We can use "tactile" simulations to give us a sense of what the computer is doing
  
 
  
- **Continuous random variables** (RVs) are more complex
  - Using the computer is the best way to simulate these

# Learning Objectives

discrete RVs simmies

## Tactile simulations
- We've already seen coin flips! 
- We can also use cards, dice, and other objects to simulate discrete random variables

 

- Other common method: A **box model** uses a box/hat/bucket of "tickets" with labels to represent possible outcomes
  - Allows us to increase the number of "tickets" with appropriate labels
  - **Coin flip as box model:** A box with two tickets (H and T).
  - **90% free throw shooter:** A box with 10 tickets (9 "make" and 1 "miss").
  - Draws can be **with replacement** (e.g., coin flips) or **without replacement** (e.g., dealing a poker hand).

## Tactile simulations: Spinners

- If draws are with replacement, we can use a single circular **spinner** instead of a box

 

- The area of each sector corresponds to the probability of an outcome

 

- Let's Google "spinner" and see what we find!
  - We can try to make a spinner for a coin flip


## Example: Dice Rolls

::: example
::: ex-title
Example: Simulating Two Rolls of a Fair Four-Sided Die
:::
::: ex-cont
Let $X$ be the sum of two rolls of a fair four-sided die, and $Y$ be the larger of the two rolls. How would we simulate $X$ and $Y$ separately?
:::
:::

## How do we extend a single simulation to useful simulations?

- We can also use `R` to sample from the box or spinner

- The `sample()` function in R allows us to simulate equally likely draws

- For example, we can simulate a coin flip
```{r}
#| echo: true

sample(x = c("H", "T"), size = 1, replace = TRUE)
```

- We can simulate our example of the four sided die
```{r}
#| echo: true

sample(x = c(1, 2, 3, 4), size = 2, replace = TRUE)
```

# Learning Objectives

continuous RVs simmies


## The Problem
- Regina and Cady's arrival times are between noon (0 minutes) and 1 PM (60 minutes).
- The arrival times are a **continuous random variable** on the interval $[0, 60]$.
- We cannot use a box model because there are uncountably many possible outcomes.
- A spinner with a continuous axis is the tactile analog.

## Uniform Model
- We can construct a spinner where the needle is equally likely to land on any value between 0 and 60.
- We would spin it twice to get Regina's and Cady's arrival times.
- 

## Normal Model
- What if Regina is more likely to arrive around 12:30?
- We can create a spinner where values near 30 are "stretched" to occupy more area.
- This is an example of a **Normal distribution**, where probabilities are concentrated around a central value.


# Learning Objectives

doing it in R

## Computer Simulation in R

- We can use R to perform simulations without a dedicated package like Symbulate.
- R has built-in functions for generating random numbers from different distributions.

## Simulating Outcomes
- The `sample()` function is a powerful tool for simulating draws from a box model.

```{r}
#| echo: true


# A box model for two rolls of a fair four-sided die
rolls <- sample(x = c(1, 2, 3, 4), size = 2, replace = TRUE)

# Simulate 10 repetitions, with each row representing a pair of rolls
reps <- 10
replicate(reps, sample(x = 1:4, size = 2, replace = TRUE))
```

- sample() can handle both with (replace = TRUE) and without (replace = FALSE) replacement.

## Simulating Random Variables

We can apply functions to the simulated outcomes to get the values of our random variables.

Let's simulate X (sum) and Y (max).

```{r}
#| echo: true


reps <- 1000
simulations <- replicate(reps, sample(x = 1:4, size = 2, replace = TRUE))

# Calculate X (the sum) for each repetition
X_simulated <- apply(simulations, 2, sum)

# Calculate Y (the max) for each repetition
Y_simulated <- apply(simulations, 2, max)

# Display the first 10 values for X and Y
head(X_simulated, 10)
head(Y_simulated, 10)
```
