---
title: "Lesson 3: Introduction to Simulations"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#006a4e"
date: "`r library(here); source(here('class_dates.R')); w2d1`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: Lesson 3 Slides
    html-math-method: mathjax
---

```{r setup, include=FALSE}
library(here)
source(here("class_dates.R"))
```

# Learning Objectives

1.  intro to simulations
2.  discrete RVs simmies
3.  continuous RVs simmies

## Recall: Outcomes, events, sample spaces

::::: definition
::: def-title
Definition: Outcome
:::

::: def-cont
The possible results in a random phenomenon.
:::
:::::

::::: definition
::: def-title
Definition: Sample Space
:::

::: def-cont
The **sample space** $S$ is the set of *all* outcomes
:::
:::::

::::: definition
::: def-title
Definition: Event
:::

::: def-cont
An **event** is a *collection of some* outcomes. An event can include multiple outcomes or no outcomes (a subset of the sample space).
:::
:::::

When thinking about events, think about outcomes that you might be asking the probability of. For example, what is the probability that you get a heads or a tails in one flip? (Answer: 1)

## We need to understand random variables 

::: definition
::: def-title
Definition: Random Variable
:::

::: def-cont
For a given sample space $S$, a **random variable** (r.v.) is a **function** whose domain is $S$ and whose range is the set of real numbers $\mathbb{R}$. **A random variable assigns a real number to each outcome in the sample space.**
:::
:::

-   A random variable's value is completely determined by the outcome $\omega$, where $\omega \in S$

    -   What is *random* is the outcome $\omega$

-   A random variable is a function from the sample space (with outcomes $\omega$) to the set of real numbers

    -   We typically write $X$ instead of $X(\omega)$, where $X$ is our random variable
    
- Thus, we can take our sample space (all outcomes) and make functional transformations to it

## The cool (and tricky) thing about random variables

Do you remember our coin example from Lesson 1? We tossed one or two coins.

- For each coin, the **sample space** is heads and tails ($S = {H,T}$)
- If we want the **sample space** for both coins in order, then we have combinations ($S={(H,H), (H,T), (T,H), (T,T)}$)

 

We make the **random variable** a *function* of the **sample space**.

- For one coin toss, we can say random variable $X$ is $1$ if we toss a heads ($\omega = \text{H}$) and $X=0$ if we get a tails
- For the two coins, we can say $X$ is the count of heads, so if $\omega = \text{(H, T)}$, $X=1$
    
## Types of random variables

There are two types of random variables:

::: columns
::: column

::: theorem
::: thm-cont
-   [**Discrete random variables** (RVs)]{style="color:#20c997;"}: the set of possible values is either finite or can be put into a countably infinite list

    -   You could *theoretically* list the specific possible outcomes that the variable can take

    -   If you sum the rolls of three dice, you must get a whole number. For example, you can't get any number between 3 and 4.
:::
:::

:::
::: column

::: axiom
::: axiom-cont
-   [**Continuous random variables** (RVs)]{style="color:#d9534f;"}: take on values from continuous *intervals*, or unions of continuous intervals

    -   Variable takes on a range of values, but there are infinitely possible values within the range

    -   If you keep track of the time you sleep, you can sleep for 8 hours or 7.9 hours or 7.99 hours or 7.999 hours ...
:::
:::
:::
:::

## What is a simulation?

A **probability model** for a random phenomenon includes a sample space, events, random variables, and a probability measure.

::::: definition
::: def-title
Simulation
:::

::: def-cont
**Simulation** involves using a probability model to artificially recreate a random phenomenon, many times, usually using a computer.
:::
:::::

We simulate outcomes and values of random variables according to the model's assumptions.

## The Foundation: Relative Frequencies

-   Probabilities can be interpreted as **long-run relative frequencies**
-   By simulating a random phenomenon a large number of times, we can approximate the probability of an event by calculating the relative frequency of its occurrence.
-   Simulation is a powerful tool for approximating probabilities, distributions of random variables, long-run averages, and other characteristics.

## Still confused about long-run relative frequencies?

[Seeing Theory, Chapter 1: Basic Probability, Chance Events](https://seeing-theory.brown.edu/basic-probability/index.html) We'll flip a coin many many times

```{css}
#| echo: false

<style>
#wrap { width: 600px; height: 390px; padding: 0; overflow: hidden; }
#frame { width: 1300px; height: 600px; border: 1px solid black; }
#frame { zoom: 1.5; -moz-transform: scale(1.5); -moz-transform-origin: 0 0; }
</style>
```

```{=html}
<iframe id="frame" src="https://seeing-theory.brown.edu/basic-probability/index.html"></iframe>
```

## Types of simulations

-   We can simulate discrete and continuous random variables

 

-   **Discrete random variables** (RVs) are a little easier to understand
    -   We can use "tactile" simulations to give us a sense of what the computer is doing

 

-   **Continuous random variables** (RVs) are more complex
    -   Using the computer is the best way to simulate these

# Learning Objectives

discrete RVs simmies

## Tactile simulations

-   We've already seen coin flips!
-   We can also use cards, dice, and other objects to simulate discrete random variables

 

-   Other common method: A **box model** uses a box/hat/bucket of "tickets" with labels to represent possible outcomes
    -   Allows us to increase the number of "tickets" with appropriate labels
    -   **Coin flip as box model:** A box with two tickets (H and T).
    -   **90% free throw shooter:** A box with 10 tickets (9 "make" and 1 "miss").
    -   Draws can be **with replacement** (e.g., coin flips) or **without replacement** (e.g., dealing a poker hand).

## Tactile simulations: Spinners

-   If draws are with replacement, we can use a single circular **spinner** instead of a box

 

-   The area of each sector corresponds to the probability of an outcome

 

-   Let's Google "spinner" and see what we find!
    -   We can try to make a spinner for a coin flip

## Example: Dice Rolls

::::: example
::: ex-title
Example: Simulating Two Rolls of a Fair Four-Sided Die
:::

::: ex-cont
We're going to roll two four-sided die. Let $X$ be the sum of two rolls, and $Y$ be the larger of the two rolls. How would we simulate $X$ and $Y$ separately?
:::
:::::

Let's build up some coding tools to do this!

## How do we simulate something like a single dice roll?

-   We can also use `R` to sample from the box or spinner

-   The `sample()` function in R allows us to simulate equally likely draws

-   For example, we can simulate a coin flip

      - What is `x`? 
      - What is `size`?

```{r}
#| echo: true

sample(x = c("H", "T"), size = 1)
```

 

- Or a dice roll

```{r}
#| echo: true

sample(x = c(1, 2, 3, 4), size = 1)
```

## What if we have multiple phenomena at once?

- We can set `size` to be larger than 1 to simulate multiple draws at once

```{r}
#| echo: true

sample(x = c("H", "T"), size = 5, replace = TRUE)
```

-   We can simulate our example of the two four-sided die

```{r}
#| echo: true

sample(x = c(1, 2, 3, 4), size = 2, replace = TRUE)
```

 

- What happens if we set `replace = FALSE`?

```{r}
#| echo: true

sample(x = c(1, 2, 3, 4), size = 2, replace = FALSE)
sample(x = c(1, 2, 3, 4), size = 4, replace = FALSE)
```

## Simulating Outcomes

-   The `sample()` function is a powerful tool for simulating draws from a box model.

```{r}
#| echo: true


# A box model for two rolls of a fair four-sided die
rolls <- sample(x = c(1, 2, 3, 4), size = 2, replace = TRUE)

# Simulate 10 repetitions, with each row representing a pair of rolls
reps <- 10
replicate(reps, sample(x = 1:4, size = 2, replace = TRUE))
```

-   sample() can handle both with (replace = TRUE) and without (replace = FALSE) replacement.

## Simulating Random Variables

We can apply functions to the simulated outcomes to get the values of our random variables.

Let's simulate X (sum) and Y (max).

```{r}
#| echo: true


reps <- 1000
simulations <- replicate(reps, sample(x = 1:4, size = 2, replace = TRUE))

# Calculate X (the sum) for each repetition
X_simulated <- apply(simulations, 2, sum)

# Calculate Y (the max) for each repetition
Y_simulated <- apply(simulations, 2, max)

# Display the first 10 values for X and Y
head(X_simulated, 10)
head(Y_simulated, 10)
```


# Learning Objectives

continuous RVs simmies

## The Problem

-   Regina and Cady's arrival times are between noon (0 minutes) and 1 PM (60 minutes).
-   The arrival times are a **continuous random variable** on the interval $[0, 60]$.
-   We cannot use a box model because there are uncountably many possible outcomes.
-   A spinner with a continuous axis is the tactile analog.

## Uniform Model

-   We can construct a spinner where the needle is equally likely to land on any value between 0 and 60.
-   We would spin it twice to get Regina's and Cady's arrival times.
-   

## Normal Model

-   What if Regina is more likely to arrive around 12:30?
-   We can create a spinner where values near 30 are "stretched" to occupy more area.
-   This is an example of a **Normal distribution**, where probabilities are concentrated around a central value.

## 4 (S)teps of a Simulation

::::: proposition
::: prop-title
1.  Set up
:::

::: prop-cont
Define the probability space and related random variables and events, including assumptions.
:::
:::::

::::: proof1
::: proof-title
2.  Simulate
:::

::: proof-cont
Run the simulation to generate outcomes according to the assumptions.
:::
:::::

::::: example
::: ex-title
3.  Summarize
:::

::: ex-cont
Analyze the output using plots and summary statistics like relative frequencies and averages.
:::
:::::

::::: extra
::: ext-title
4.  Sensitivity analysis
:::

::: ext-cont
Investigate how results change when assumptions or parameters of the model are altered.
:::
:::::
